PLANETESIMAL ACCRETION AT SHORT ORBITAL PERIODS

### 2022-08-04 main.tex
Full read through.

Planetesimal accretion is a key phase in the terrestrial planet growth process, bridging the gap from kilometer-sized bodies up to roughly moon-sized objects known as planetary embryos. In the earliest stages of the planet formation process, aerodynamic forces dominate the growth and evolution of the solids and statistical models \citep{johansen14, birnstiel16} are appropriate to describe how these numerous, small bodies coagulate. Due to the internal pressure support of the gas disk, the gas itself orbits at sub-Keplerian speed and exerts a headwind on any solids large enough to decouple from the gas \citep{weidenschilling77}. Above about a meter in size, this headwind is maximally effective at sapping away angular momentum and planet-building material can fall onto the central star on catastrophically short timescales \citep{weidenschilling77, nakagawa86}. Additionally, laboratory experiments suggest that collisions between mm- to cm- sized solids tend to result in bouncing or destruction, rather than continued growth \citep{blum93, beitz11, colwell03}. For these reasons, a number of mechanisms which involve radially concentrating solids in a planet-forming disk have been proposed to facilitate fast growth from mm to km sizes \citep{johansen07, lyra08, bai10}. Interestingly, formation models for the short-period multiplanet systems revealed by Kepler \citep{fabrycky14} also seem to require enhanced concentrations of planet-building material to reproduce the observed architectures \citep{raymond07, hansen12}.

@@

After a long period of quiescence, the collection of embryos and remaining planetesimals undergoes a large-scale instability \citep{chambers98}. Suddenly, the oligarchics are no longer on isolated, stable orbits and coalesce to form Earth-sized planets through a series of extremely energetic, giant impacts \citep{kokubo02, raymond05, raymond06}. Due to the relative ease of modeling the early dust coagulation phases and the final giant impact phase, these steps in the terrestrial planet formation process have received the most attention in the literature. The planetesimal accretion phase, which we will focus on in this paper, falls into an awkward in-between, where there are too many particles to directly track with traditional N-body codes, while the gravitational influence of the few, but massive oligarchs that form make statistical methods an inappropriate approach. Because of this limitation, planetesimal accretion is usually modeled in a narrow ring and the results are then scaled to suit whatever situation is being studied. N-body simulations of terrestrial planet formation typically begin with a series of neatly spaced oligarchs, whose mass varies smoothly with orbital period. As we will show in this paper, this extrapolation with orbital period is inappropriate, particularly at short ($<$ 10 day) periods.

Given that systems of tightly-packed inner planets (STIPs) appear to be a common outcome of planet formation, understanding exactly how solids accumulate at short orbital periods is crucial. Although gas-disk driven migration of the planets themselves is often invoked to explain the observed architectures, we will focus on an in-situ model in this paper. That is, once the planetesimals themselves form, they largely stay in place and any subsequent large-scale movement of the solids are the result of mutual gravitational interactions. The focus of this work will be to understand how the outcome of the planetesimal accretion process scales with orbital period by using a high-powered N-body code to directly follow the growth and evolution of the planetesimals across a wide range of orbital periods (1 to 100 days). In doing so, we will assess whether the typical `isolation mass' initial conditions used in studies of terrestrial planet formation are actually appropriate for understanding STIPs. In a series of follow up papers, we plan to use these results to directly track the formation of full-sized planets from planetesimals and to understand how gravitational perturbations from a nearby gas giant might affect the outcome of this process.

% This stuff doesn't fit since I reworked the intro. Might be better to save some of this for paper 2 (planetesimals->planets)
% Still deciding whether I should mention habitability and low mass stars
% I think there are plenty of STIPs around sun like stars, but I should double check this

@@

To investigate whether any of these planetesimal disks underwent
runaway growth, we examine the time evolution of the maximum and mean
masses in each case. The ratio $m_{max}/\left< m \right>$ is plotted
in figure \ref{fig:alpha_beta_evo}. On this plot, a positive slope
indicates that the growth rate is accelerating with mass, which is
evident at various times in all four cases. Even with a large value of
$\beta$, which means that the effective collision cross section is
very near the geometric value, runaway growth still appears to
operate. This seems to imply that, as bodies collide and grow, the
relative increase in their gravitational focusing factors, rather than
their absolute size, is what drives the system towards runaway
growth. Although larger encounter velocities will lengthen the growth
timescales, this mode of growth appears to be inevitable, so long as
gravity is the dominant force in the system.
% TRQ: also mention the negative slope

@@

\noindent where $\Sigma_{g}$ is the gas surface density (taken to be 240x the solid surface density), $h_{g} = c_{s} / \Omega$ is the local gas scale height and $z$ is the height above the disk midplane. The sound speed profile is given by $c_{s} = \sqrt{k_{B} T(r) / \left( \mu m_{H} \right)}$, where $k_{B}$ is Boltzmann's constant, $T(r) = T_{0} r^{-Q}$, $\mu = 2.34$ and $m_{h}$ is the mass of a hydrogen atom. For a protoplanetary disk around a typical M star, $T_{0} = 148$ K and $Q$ = 0.58 \citep{andrews05}.

@@

\noindent where $v_{k}$ is the local Keplerian velocity. As in section \ref{sec:narrow}, the argument of perihelion $\omega$, longitude of ascending node $\Omega$, and mean anomaly M for the planetesimals drawn from a uniform distribution $\in [0, 2 \pi)$.

@@

In the case of the fdHi simulation, there are nearly 1 million particles, whose orbital periods two orders of magnitude in time. Due to the very short interaction timescales near the inner edge of the disk, a fixed timestep size would required a prohibitively large number of steps to follow planetesimal growth throughout the entire disk. For this reason, we use a multi-tiered timestepping scheme, in which particles are placed onto the nearest power of two timestep based on their most recently calculated gravitational acceleration. (Mention that this scheme is used in other works with ChaNGa).

This more efficient scheme, however, introduces two additional issues related to particles interacting on different timesteps. First, momentum is not completely conserved when bodies switch timesteps. The error introduced becomes particularly severe for a particle on an eccentric Keplerian orbit, whose perihelion and aphelion distances straddle a timestep boundary. For a large collection of particles, this problem manifests itself as a V-shaped gap in the semimajor axis-eccentricity plane, which is centered on the boundary itself. To correct this problem, we introduce a second timestepping criterion, which is based on the gravitational acceleration on the particle by the central star at perihelion. Only in the case of a close encounter with another planetesimal (in which the acceleration is no longer dominated by the star) is the timestep allowed to reduce based on the original criterion.

@@

\begin{table}
\caption{Final Properties of Full Disk Simulations}
\begin{tabular}{lllll} \hline \hline
Name     & $M_{PP}^a$ & $T_{int}^b$ & $T_{int 1}^c$ &  \\ \hline
fdHi                 & 100                                                         & 1.5                              & 4  &  \\
fdHiShallow     & 100                                                        & 0.5                              & 4  &  \\
fdHiSteep       & 100                                                         & 2.5                              & 4 &  \\
fdLo                & 1                                                             & 1.5                              & 4 &  \\ \hline
\end{tabular}
\label{tab:sim_properties}
\begin{flushleft}
\textsc{Note} \\ {$^a$ Maximum protoplanet mass [$M_{earth}$] \\
		       $^b$ Integration time [yr] \\
		       $^c$ Rescaled integration time [yr]} \\
\end{flushleft}
\end{table}

In all four simulations, the timescales for embryo formation depend on the chosen surface density profile, along with the particular location in the disk. Protoplanets form first near the inner edge of the disk, where the dynamical timescales are short. Growth proceeds in an inside-out fashion, with the outermost regions of the disk completing the protoplanet assembly phase last. This fact is not typically accounted for in planet formation simulations, and appears to be an important component to forming realistic solar system analogs \citep{clement20}. As with the narrow annulus simulations, we stop the integration once the mass of protoplanets in the outermost region of the disk reach a steady value. In table \ref{tab:sim_properties}, we summarize the outcomes of the four `full disk' cases.

We show the final state of the `fiHi' simulation in figure \ref{fig:fulldisk_e_m}. In the top panel, the initial (contours) and final (points) state of the simulation is shown in orbital period and eccentricity. The size of the points indicates the relative mass of the bodies. In the bottom panel, the mass of the largest bodies is shown as a function of orbital period. The solid curves indicate the isolation mass

\begin{equation}\label{eq:iso}
	M_{iso} = \left[ \frac{\left( 2 \pi a^2 \Sigma \tilde{b} \right)^3}{3 M_{*}} \right]^{1/2},
\end{equation}

\noindent where $\tilde{b}$ is the size of the feeding zone in units of Hill radii. For a body on a circular orbit $\tilde{b} = 2 \sqrt{3}$, while a protoplanet formed by oligarchic growth typically has $\tilde{b} = 10$ \citep{kokubo98}. The isolation mass is plotted in blue and orange for these two values of $\tilde{b}$, respectively.

@@

To ensure that the boundary seen near 60 days in orbital period is not simply a transient product of the inside out growth throughout the disk, we examine the time evolution of $\sigma/\Sigma$ at multiple locations. In figure \ref{fig:pl_frac_time}, the value of $\sigma/\Sigma$ is plotted as a function of time in 10 orbital period bins. Bins interior to 60 days are indicated with a dashed line, while those exterior are indicated with a solid line. The planetesimal surface density fraction has decelerated and is approaching an asymptotic value in all cases. In the inner disk, this is driven by the near-total depletion of the planetesimal reservoir. In the outer disk, dynamical friction between the embryos and planetesimals throttles subsequent accretion.

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/fulldisk_e_m.png}
    \caption{The final state of the fdHi simulation. In the top panel, the contours denote the initial period-eccentricity distribution of the planetesimals. Point sizes indicate the relative masses of bodies. In the bottom panel, the solid line indicates the isolation mass for $b = 2 \sqrt{3}$. The vertical dashed line marks the approximate location of the accretion mode boundary. For $P_{orb} < 60$ days, the high $\alpha$ accretion mode produces embryos that exceed the isolation mass.\label{fig:fulldisk_e_m}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/surfden_iso.png}
    \caption{Masses of the remaining bodies from the simulations listed in table \ref{tab:sim_properties}, in units of the isolation mass for $b = 2 \sqrt{3}$. Despite the vastly different initial solid surface density profiles, the orbital period inside of which the embryo masses exceed the isolation mass is very consistent.\label{fig:surfden_iso}}
\end{center}
\end{figure}

@@

Next, we compare the effect of varying the initial solid surface density profile on the resulting planetesimal and embryo distribution. The final orbital period-eccentricity state of the particles is shown in figure \ref{fig:surfden_profiles}, with point sizes indicating the relative masses of the bodies. There is a clear trend between solid surface density at a given orbital period and the eccentricities of the remaining planetesimals. This trend is largely set by the initial conditions, as the gravitational stirring is more vigorous when bodies are packed together closely and higher eccentricities are required to produce an equilibrating aerodynamic drag force (see equations \ref{eq:vs_timescale} and \ref{eq:ts_stokes}). Despite having significantly different masses, the planetary embryos formed in each case have remarkably similar eccentricities. This is likely due to the fact that inelastic collisions play a more significant role where the solid surface density is highest, which offsets the fact that the initial bodies started off in a dynamically hotter state.

@@

We begin by investigating the `smoothness' of the accretion events that give rise to each embryo. Drawing from a common technique used for cosmological simulations (citation?), we divide growth events for a given body into `major' and `minor' mergers. Minor events are defined as any collision involving an initial planetesimal, while major events consist of any two larger bodies. In figure \ref{fig:minor_frac}, we show the total mass fraction gained by minor events as a function of mass for the 100 largest bodies at the end of fdHi. Blue dots represent bodies interior to the 60 day orbital period boundary, while red crosses indicate bodies exterior. For the outer bodies in the oligarchic growth region, we see a very weak trend between minor accretion mass fraction and total mass. Regardless of how much an embryo has grown, the initial mass planetesimals contribute about the same relative amount to its total mass. At shorter orbital periods, however, there is a somewhat stronger trend. The largest bodies in this region gain about an order of magnitude more of their mass fraction from major mergers, compared to the smallest bodies.

@@

The difference in slope of the short vs longer period bodies in figure \ref{fig:minor_frac} indicates a difference in how the planetesimal-embryo populations interact in each region. In the long period oligarchic growth region, embryos continue to accrete planetesimals as they grow and the minor mass fraction stays mostly flat. At short period, collisions mostly occur between equal mass bodies as the embryos grow and the contribution from planetesimals becomes less significant over time. This suggests that embryo-embryo collisions are common at short period. As discussed in section \ref{sec:sizeandhill}, the importance of gravitational scattering is greatly reduced here. Without frequent gravitational interactions between the embryos and planetesimals, orbital repulsion \citep{kokubo98} and isolation of the embryos does not occur. As we showed in figure \ref{fig:surfden_iso}, the embryos formed in this region exceed the isolation mass.

Further evidence for the lack of gravitational scattering events can be seen in figure \ref{fig:acc_zones}. Here, we have chosen the 20 most massive bodies from the fdHi simulation and arranged them in order of present orbital period. The vertical black line indicates the orbital period of each embryo. The associated histogram for each body denotes where the planetesimals used to construct that embryo originated in the disk. The peaks of the histograms have all been normalized to the same value, so that the distributions only show the relative widths and locations of the feeding zones. Beyond the 60 day boundary, the ordering of the feeding zones and the embryo locations become much less consistent. This suggests that scattering events between embryos are much more common beyond this accretion mode boundary. In the inner region of the disk, close encounters between embryos often result in a merger, which allows them to grow beyond the isolation mass while only a minor alteration to their orbital period. With few close encounters resulting in a scattering event, embryos are unable to move away from their birth location.

@@

In all cases shown so far, the boundary between the standard oligarchic growth and the highly-efficient short period accretion region lies around 60 days in orbital period. The mode of accretion is set entirely by the local value of $\alpha$, which scales with both distance from the star and the bulk density of the planetesimals (see equation \ref{eq:alpha}). The artificial inflation of the collision cross section of the particles in our simulations, which is meant to reduce the computational expense, has the side effect of reducing the bulk densities. Because $\alpha \sim \rho_{pl}^{-1/3}$, increasing the particle sizes by a factor of f=6 reduces the orbital period at which $\alpha$ has a given value by a factor of approximately 15. Therefore, one would expect the accretion boundary to lie near 5 days in orbital period for 3 g $cm^{-3}$ bodies.

Although a simulation with f=1 is not computationally tractable, we can test whether the accretion boundary moves in the way we expect by modestly changing the value of $f$. In figure \ref{fig:f6f4}, we compare the fdHi simulation to a nearly identical run using f=4. In the top panel, the embryo masses in units of the isolation mass are shown as a function of orbital period. In the bottom panel, the value of $\alpha$ as a function of orbital period is shown for 3 g $cm^{-3}$ bodies with an artificial radius enhancement of f=1, 4 and 6. The horizontal dashed line indicates the value of alpha below which the accretion mode switches to oligarchic growth. Comparing the top and bottom panels, the intersection of the embryo masses and local isolation mass matches well with the orbital period at which $\alpha \sim 0.1$. Also shown by the curved dotted lines are the $\alpha$ curves for realistic-sized bodies with $\rho_{pl}$ = 1, 5 and 10 g $cm^{-3}$. In all cases, the accretion boundary still lies within the part of the disk in which solids would be expected to accumulate (citation?).

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/f6f4.png}
    \caption{In the top panel, we show the final masses of the bodies from the fdHi and fdHif4 simulations in units of the isolation mass for $b = 2\sqrt{3}$. The bottom panel shows the variation of $\alpha$ with orbital period for the bodies used in each case (solid curves). The orbital period at which $\alpha \simeq 0.1$ matches well with the intersection of the embryo mass distribution and isolation mass. The shaded region in the bottom panel show the values of alpha for realistic-sized (f=1) planetesimals with bulk densities between 10 and 1 g cm$^{-3}$.\label{fig:f6f4}}
\end{center}
\end{figure}

@@

To test whether a more restrictive collision model should modify our results, we present a smaller scale test using a slightly more sophisticated collision model. In this case, a collision can result in one of two outcomes: if the impact velocity is smaller than the mutual escape velocity of the colliding particles, defined as

%TRQ: I corrected this off the top of my head: CHECK!
\begin{equation}\label{eq:v_mut}
	v_{mut, esc} = \sqrt{\frac{G (m_{1} + m_{2})}{r_{1} + r_{2}}},
\end{equation}

\noindent where $m_{1}, m_{2}$ and $r_{1}, r_{2}$ are the masses and radii of colliding particles 1 and 2, then the bodies perfectly merge. For impact velocities larger than $v_{mut, esc}$, no mass is transferred and the bodies undergo a completely elastic bounce. Because there is no possibility of partial accretion, this model likely restricts growth more than the model used by \citet{leinhardt12}. However, we will show below that the bounce-merge model does not meaningfully affect the outcome of the planetesimal accretion phase, and so a partial accretion model should do the same.

@@

Radial mixing appears to be enhanced by a partial collision model, along with smaller collision cross section inflation \citep{childs22}.

@@

What is the migration timescale for the embryos to move to the inner edge of the disk? Maybe this section isn't terribly important and should be excluded?

@@

To connect these accretion modes to the final orbital architecture, and to ultimate determine what implications an in situ formation model has for the growth of STIPs, we will continue to evolve the final simulation snaphots presented here with a hybrid symplectic integrator. The final distribution of planets formed, along with composition predictions generated by applying cosmochemical models to our initial planetesimal distributions and propagating compositions through the merger trees, will be examined in a follow-up paper.

>>>>

Planetesimal accretion is a key phase in the terrestrial planet growth
process, bridging the gap from kilometer-sized bodies up to roughly
moon-sized objects known as planetary embryos. In the earliest stages
of the planet formation process, aerodynamic forces dominate the
growth and evolution of the solids and statistical models
\citep{johansen14, birnstiel16} are appropriate to describe how these
numerous, small bodies coagulate. Due to the internal pressure support
of the gas disk, the gas itself orbits at sub-Keplerian speed and
exerts a headwind on any solids large enough to decouple from the gas
\citep{weidenschilling77}. Above about a meter in size, this headwind
is maximally effective at sapping away angular momentum, and planet-building material can fall onto the central star on catastrophically short timescales \citep{weidenschilling77, nakagawa86}. Additionally, laboratory experiments suggest that collisions between mm- to cm- sized solids tend to result in bouncing or destruction, rather than continued growth \citep{blum93, beitz11, colwell03}. For these reasons, a number of mechanisms which involve radially concentrating solids in a planet-forming disk have been proposed to facilitate fast growth from mm to km sizes \citep{johansen07, lyra08, bai10}. Interestingly, formation models for the short-period multiplanet systems revealed by Kepler \citep{fabrycky14} also seem to require enhanced concentrations of planet-building material to reproduce the observed architectures \citep{raymond07, hansen12}.

@@

After a long period of quiescence, the collection of embryos and
remaining planetesimals undergoes a large-scale instability
\citep{chambers98}.
%TRQ "Suddenly" seems a drastic term to use here.
Suddenly, the oligarchics are no longer on isolated, stable orbits and
coalesce to form Earth-sized planets through a series of extremely
energetic, giant impacts \citep{kokubo02, raymond05, raymond06}.
%TRQ new paragraph here?
Due to the relative ease of modeling the early dust coagulation phases
and the final giant impact phase, these steps in the terrestrial
planet formation process have received the most attention in the
literature. The planetesimal accretion phase, which we will focus on
in this paper, falls into an awkward in-between, where there are too
many particles to directly track with traditional N-body codes, while
the gravitational influence of the few, but massive oligarchs that
form make statistical methods an inappropriate approach. Because of
this limitation, planetesimal accretion is usually modeled in a narrow
ring, and the results are then scaled to suit whatever situation is being studied. N-body simulations of terrestrial planet formation typically begin with a series of neatly spaced oligarchs, whose mass varies smoothly with orbital period. As we will show in this paper, this extrapolation with orbital period is inappropriate, particularly at short ($<$ 10 day) periods.

Given that systems of tightly-packed inner planets (STIPs) appear to
be a common outcome of planet formation, understanding exactly how
solids accumulate at short orbital periods is crucial. Although
gas-disk driven migration of the planets themselves is often invoked
to explain the observed architectures, we will focus on an in-situ
model in this paper. That is, once the planetesimals themselves form,
they largely stay in place and any subsequent large-scale movement of
the solids are the result of mutual gravitational interactions. The
focus of this work will be to understand how the outcome of the
planetesimal accretion process scales with orbital period by using a
high-powered N-body code to directly follow the growth and evolution
of the planetesimals across a wide range of orbital periods (1 to 100
days). In doing so, we will assess whether the typical ``isolation
mass'' initial conditions used in studies of terrestrial planet formation are actually appropriate for understanding STIPs. In a series of follow up papers, we plan to use these results to directly track the formation of full-sized planets from planetesimals and to understand how gravitational perturbations from a nearby gas giant might affect the outcome of this process.

% This stuff doesn't fit since I reworked the intro. Might be better to save some of this for paper 2 (planetesimals->planets)
% Still deciding whether I should mention habitability and low mass stars
% I think there are plenty of STIPs around sun like stars, but I should double check this
% TRQ: for the discussion/conclusions: mention that the mass of the
% star doesn't matter (caveat composition changes in the
% planetesimals); it's the orbital period that is crucial.

@@

To investigate whether any of these planetesimal disks underwent
runaway growth, we examine the time evolution of the maximum and mean
masses in each case. The ratio $m_{max}/\left< m \right>$ is plotted
in figure \ref{fig:alpha_beta_evo}. On this plot, a positive slope
indicates that the growth rate is accelerating with mass, which is
evident at various times in all four cases. Even with a large value of
$\beta$, which means that the effective collision cross section is
very near the geometric value, runaway growth still appears to
operate. This seems to imply that, as bodies collide and grow, the
relative increase in their gravitational focusing factors, rather than
their absolute size, is what drives the system towards runaway
growth. Although larger encounter velocities will lengthen the growth
timescales, this mode of growth appears to be inevitable, so long as
gravity is the dominant force in the system.
% TRQ: also mention the negative slope?

@@

\noindent where $\Sigma_{g}$ is the gas surface density (taken to be
240 times the solid surface density), $h_{g} = c_{s} / \Omega$ is the local gas scale height and $z$ is the height above the disk midplane. The sound speed profile is given by $c_{s} = \sqrt{k_{B} T(r) / \left( \mu m_{H} \right)}$, where $k_{B}$ is Boltzmann's constant, $T(r) = T_{0} r^{-Q}$, $\mu = 2.34$ and $m_{h}$ is the mass of a hydrogen atom. For a protoplanetary disk around a typical M star, $T_{0} = 148$ K and $Q$ = 0.58 \citep{andrews05}.

@@

\noindent where $v_{k}$ is the local Keplerian velocity. As in section
\ref{sec:narrow}, the argument of perihelion $\omega$, longitude of
ascending node $\Omega$, and mean anomaly M for the planetesimals are drawn from a uniform distribution $\in [0, 2 \pi)$.

@@

In the case of the fdHi simulation, there are nearly 1 million
particles, whose orbital periods vary by two orders of magnitude. Due
to the very short interaction timescales near the inner edge of the
disk, a fixed timestep size would required a prohibitively large
number of steps to follow planetesimal growth throughout the entire
disk. For this reason, we use a multi-tiered timestepping scheme, in
which particles are placed onto the nearest power of two timestep
based on their most recently calculated gravitational
acceleration. (This scheme is used on almost all works using ChaNGa,
and is common among large-scale simulation codes.)

This more efficient scheme, however, introduces two additional issues
related to particles interacting on different timesteps. First,
momentum is not completely conserved when bodies switch timesteps. The
error introduced becomes particularly severe for a particle on an
eccentric Keplerian orbit, whose perihelion and aphelion distances
straddle a timestep boundary. For a large collection of particles,
this problem manifests itself as the development of a V-shaped gap in the semimajor axis-eccentricity plane, which is centered on the boundary itself. To correct this problem, we introduce a second timestepping criterion, which is based on the gravitational acceleration on the particle by the central star at perihelion. Only in the case of a close encounter with another planetesimal (in which the acceleration is no longer dominated by the star) is the timestep allowed to reduce based on the original criterion.

@@

\begin{table}
\caption{Final Properties of Full Disk Simulations}
%TRQ: the mass numbers in this table look wrong.
\begin{tabular}{lllll} \hline \hline
Name     & $M_{PP}^a$ & $T_{int}^b$ & $T_{int 1}^c$ &  \\ \hline
fdHi                 & 100                                                         & 1.5                              & 4  &  \\
fdHiShallow     & 100                                                        & 0.5                              & 4  &  \\
fdHiSteep       & 100                                                         & 2.5                              & 4 &  \\
fdLo                & 1                                                             & 1.5                              & 4 &  \\ \hline
\end{tabular}
\label{tab:sim_properties}
\begin{flushleft}
\textsc{Note} \\ {$^a$ Maximum protoplanet mass [$M_{earth}$] \\
		       $^b$ Integration time [yr] \\
		       $^c$ Rescaled integration time [yr]} \\
\end{flushleft}
\end{table}

In all four simulations, the timescales for embryo formation depend on
the chosen surface density profile, along with the particular location
in the disk. Protoplanets form first near the inner edge of the disk,
where the dynamical timescales are short. Growth proceeds in an
inside-out fashion, with the outermost regions of the disk completing
the protoplanet assembly phase last. This fact is not typically
accounted for in planet formation simulations, and appears to be an
important component to forming realistic solar system analogs
\citep{clement20}. As with the narrow annulus simulations, we stop the
integration once the masses of protoplanets in the outermost region of
the disk reach a steady value. In table \ref{tab:sim_properties}, we
summarize the outcomes of the four ``full disk'' cases.

We show the final state of the ``fiHi'' simulation in figure \ref{fig:fulldisk_e_m}. In the top panel, the initial (contours) and final (points) state of the simulation is shown in orbital period and eccentricity. The size of the points indicates the relative mass of the bodies. In the bottom panel, the mass of the largest bodies is shown as a function of orbital period. The solid curves indicate the isolation mass,
\begin{equation}\label{eq:iso}
	M_{iso} = \left[ \frac{\left( 2 \pi a^2 \Sigma \tilde{b} \right)^3}{3 M_{*}} \right]^{1/2},
\end{equation}
where $\tilde{b}$ is the size of the feeding zone in units
of Hill radii. For a body on a circular orbit $\tilde{b} = 2
\sqrt{3}$, while a protoplanet formed by oligarchic growth typically
has $\tilde{b} = 10$ \citep{kokubo98}.
%TRQ there is no orange line in the figure.
The isolation mass is plotted in blue and orange for these two values of $\tilde{b}$, respectively.

@@

To ensure that the boundary seen near 60 days in orbital period is not
simply a transient product of the inside out growth throughout the
disk, we examine the time evolution of $\sigma/\Sigma$ at multiple
locations. In figure \ref{fig:pl_frac_time}, the value of
$\sigma/\Sigma$ is plotted as a function of time in 10 orbital period
bins.
%TRQ: width of the bins?
Bins interior to 60 days are indicated with a dashed line, while those exterior are indicated with a solid line. The planetesimal surface density fraction has decelerated and is approaching an asymptotic value in all cases. In the inner disk, this is driven by the near-total depletion of the planetesimal reservoir. In the outer disk, dynamical friction between the embryos and planetesimals throttles subsequent accretion.

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/fulldisk_e_m.png}
    \caption{The final state of the fdHi simulation. In the top panel,
      the contours denote the initial period-eccentricity distribution
      of the planetesimals. Point sizes indicate the relative masses
      of bodies. In the bottom panel, the solid line indicates the
      isolation mass for $\tilde{b} = 2 \sqrt{3}$. The vertical dashed line marks the approximate location of the accretion mode boundary. For $P_{orb} < 60$ days, the high $\alpha$ accretion mode produces embryos that exceed the isolation mass.\label{fig:fulldisk_e_m}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/surfden_iso.png}
    \caption{Masses of the remaining bodies from the simulations
      listed in table \ref{tab:sim_properties}, in units of the
      isolation mass for $\tilde{b} = 2 \sqrt{3}$. Despite the vastly different initial solid surface density profiles, the orbital period inside of which the embryo masses exceed the isolation mass is very consistent.\label{fig:surfden_iso}}
\end{center}
\end{figure}

@@

Next, we compare the effect of varying the initial solid surface
density profile on the resulting planetesimal and embryo
distribution. The final orbital period-eccentricity state of the
particles is shown in figure \ref{fig:surfden_profiles}, with colors
and point styles indicating the different initial surface density profiles,
and with point sizes indicating the relative masses of the bodies. There is a clear trend between solid surface density at a given orbital period and the eccentricities of the remaining planetesimals. This trend is largely set by the initial conditions, as the gravitational stirring is more vigorous when bodies are packed together closely and higher eccentricities are required to produce an equilibrating aerodynamic drag force (see equations \ref{eq:vs_timescale} and \ref{eq:ts_stokes}). Despite having significantly different masses, the planetary embryos formed in each case have remarkably similar eccentricities. This is likely due to the fact that inelastic collisions play a more significant role where the solid surface density is highest, which offsets the fact that the initial bodies started off in a dynamically hotter state.

@@

We begin by investigating the ``smoothness'' of the accretion events
that give rise to each embryo. Drawing from a common technique used
for cosmological simulations (citation?), we divide growth events for
a given body into ``major'' and ``minor'' mergers. Minor events are defined as any collision involving an initial planetesimal, while major events consist of any two larger bodies. In figure \ref{fig:minor_frac}, we show the total mass fraction gained by minor events as a function of mass for the 100 largest bodies at the end of fdHi. Blue dots represent bodies interior to the 60 day orbital period boundary, while red crosses indicate bodies exterior. For the outer bodies in the oligarchic growth region, we see a very weak trend between minor accretion mass fraction and total mass. Regardless of how much an embryo has grown, the initial mass planetesimals contribute about the same relative amount to its total mass. At shorter orbital periods, however, there is a somewhat stronger trend. The largest bodies in this region gain about an order of magnitude more of their mass fraction from major mergers, compared to the smallest bodies.

@@

The difference in slope of the short vs.\ long period bodies in figure \ref{fig:minor_frac} indicates a difference in how the planetesimal-embryo populations interact in each region. In the long period oligarchic growth region, embryos continue to accrete planetesimals as they grow and the minor mass fraction stays mostly flat. At short period, collisions mostly occur between equal mass bodies as the embryos grow and the contribution from planetesimals becomes less significant over time. This suggests that embryo-embryo collisions are common at short period. As discussed in section \ref{sec:sizeandhill}, the importance of gravitational scattering is greatly reduced here. Without frequent gravitational interactions between the embryos and planetesimals, orbital repulsion \citep{kokubo98} and isolation of the embryos does not occur. As we showed in figure \ref{fig:surfden_iso}, the embryos formed in this region exceed the isolation mass.

Further evidence for the lack of gravitational scattering events in
the inner region can be seen in figure \ref{fig:acc_zones}. Here, we
have chosen the 20 most massive bodies from the fdHi simulation and
arranged them in order of present orbital period. The vertical black
line indicates the orbital period of each embryo. The associated
histogram for each body denotes where the planetesimals used to
construct that embryo originated in the disk. The peaks of the
histograms have all been normalized to the same value, so that the
distributions only show the relative widths and locations of the
feeding zones. Beyond the 60 day boundary, the ordering of the feeding
zones and the embryo locations become much less consistent. This
suggests that scattering events between embryos are much more common
beyond this accretion mode boundary. In the inner region of the disk,
close encounters between embryos often result in a merger, which
allows them to grow beyond the isolation mass with only a minor alteration to their orbital period. With few close encounters resulting in a scattering event, embryos are unable to move away from their birth location.

@@

In all cases shown so far, the boundary between the standard
oligarchic growth and the highly-efficient short period accretion
region lies around 60 days in orbital period. The mode of accretion is
set entirely by the local value of $\alpha$, which scales with both
distance from the star and the bulk density of the planetesimals (see
equation \ref{eq:alpha}). The artificial inflation of the collision
cross section of the particles in our simulations, which is meant to
reduce the computational expense, has the side effect of reducing the
bulk densities. Because $\alpha \sim \rho_{pl}^{-1/3}$, increasing the
particle sizes by a factor of $f=6$ reduces the orbital period at which $\alpha$ has a given value by a factor of approximately 15. Therefore, one would expect the accretion boundary to lie near 5 days in orbital period for 3 g $cm^{-3}$ bodies.

Although a simulation with $f=1$ is not computationally tractable, we
can test whether the accretion boundary moves in the way we expect by
modestly changing the value of $f$. In figure \ref{fig:f6f4}, we
compare the fdHi simulation to a nearly identical run using $f=4$. In
the top panel, the embryo masses in units of the isolation mass are
shown as a function of orbital period. In the bottom panel, the value
of $\alpha$ as a function of orbital period is shown for 3 g $cm^{-3}$
bodies with an artificial radius enhancement of $f=1$, 4 and 6. The
horizontal dashed line indicates the value of alpha below which the
accretion mode switches to oligarchic growth. Comparing the top and
bottom panels, the intersection of the embryo masses and local
isolation mass matches well with the orbital period at which $\alpha
\sim 0.1$.  Also shown by the shaded region are $\alpha$ values for
realistic sized bodies with $\rho_{pl}$ between 1 and 10 g cm$^{-3}$.
% Also shown by the curved dotted lines are the $\alpha$ curves for
% realistic-sized bodies with $\rho_{pl}$ = 1, 5 and 10 g $cm^{-3}$.
In all cases, the accretion boundary still lies within the part of the disk in which solids would be expected to accumulate (citation?).

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/f6f4.png}
    \caption{In the top panel, we show the final masses of the bodies
      from the fdHi and fdHif4 simulations in units of the isolation
      mass for $\tilde{b} = 2\sqrt{3}$. The bottom panel shows the variation of $\alpha$ with orbital period for the bodies used in each case (solid curves). The orbital period at which $\alpha \simeq 0.1$ matches well with the intersection of the embryo mass distribution and isolation mass. The shaded region in the bottom panel show the values of alpha for realistic-sized ($f=1$) planetesimals with bulk densities between 10 and 1 g cm$^{-3}$.\label{fig:f6f4}}
\end{center}
\end{figure}

@@

To test whether a more restrictive collision model should modify our results, we present a smaller scale test using a slightly more sophisticated collision model. In this case, a collision can result in one of two outcomes: if the impact velocity is smaller than the mutual escape velocity of the colliding particles, defined as
%TRQ: I corrected this off the top of my head: CHECK!
\begin{equation}\label{eq:v_mut}
	v_{mut, esc} = \sqrt{\frac{G (m_{1} + m_{2})}{r_{1} + r_{2}}},
\end{equation}
where $m_{1}, m_{2}$ and $r_{1}, r_{2}$ are the masses and radii of
colliding particles 1 and 2, then the bodies perfectly merge. For
impact velocities larger than $v_{mut, esc}$, no mass is transferred, and the bodies undergo a completely elastic bounce. Because there is no possibility of partial accretion, this model likely restricts growth more than the model used by \citet{leinhardt12}. However, we will show below that the bounce-merge model does not meaningfully affect the outcome of the planetesimal accretion phase, and so a partial accretion model should do the same.

@@

%TRQ: expand on this?  Compare figure 10 histograms for the two runs?
Radial mixing appears to be enhanced by a partial collision model, along with smaller collision cross section inflation \citep{childs22}.

@@

What is the migration timescale for the embryos to move to the inner edge of the disk? Maybe this section isn't terribly important and should be excluded?
TRQ: I agree

@@

To connect these accretion modes to the final orbital architecture,
and to ultimately determine what implications an in situ formation model has for the growth of STIPs, we will continue to evolve the final simulation snaphots presented here with a hybrid symplectic integrator. The final distribution of planets formed, along with composition predictions generated by applying cosmochemical models to our initial planetesimal distributions and propagating compositions through the merger trees, will be examined in a follow-up paper.

\section{Acknowledgements}

NSF, XSEDE.
